Main Analysis

### 2024-05-29 Manuscript Analysis/adaptation_analysis.Rmd
updated figures

rep_acc_df |>
  dplyr::mutate(trial_c = trial_c + 1) |>
  dplyr::group_by(speaker, trial_c, rep_acc) |>
  dplyr::summarize(n = n()) |>
  dplyr::ungroup() |>
  tidyr::pivot_wider(names_from = "rep_acc",
                     values_from = "n") |>
  dplyr::mutate(accurate = ifelse(is.na(accurate), 0, accurate),
                inaccurate = ifelse(is.na(inaccurate), 0, inaccurate),
                prop_accurate = accurate/(accurate + inaccurate)) |>
  ggplot() +
  aes(x = trial_c,
      y = prop_accurate,
      group = speaker,
      color = speaker,
      fill = speaker, 
      linetype = speaker,
      shape = speaker) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(y = "Proportion of Accurate Responses",
       x = "Trial Number") +
  theme_bw() +
  theme(legend.position = "bottom",
        aspect.ratio = 1) +
  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
  scale_fill_manual(values = c(my_pal[1], my_pal[5])) +
  scale_shape_manual(values = c(1, 16)) +
  scale_linetype_manual(values = c("solid", "dashed"))

@@

ple_data |>
  dplyr::mutate(trial_c = trial_c + 1) |>
  dplyr::group_by(speaker, rep_acc, trial_c) |>
  dplyr::summarize(per_effort = mean(effort_rating, na.rm = T), 
                   sd = sd(effort_rating, na.rm = T),
                   se = sd/sqrt(n())) |>
  ggplot() +
   aes(x = trial_c,
       y = per_effort,
       group = speaker,
       color = speaker,
       fill = speaker,
       shape = speaker,
       linetype = speaker) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Trial", y = "Perceived Listening Effort Rating") +
  facet_wrap("rep_acc") +
  theme_bw() +
  theme(legend.position = "bottom",
        aspect.ratio = 1) +
  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
  scale_fill_manual(values = c(my_pal[1], my_pal[5])) +
  scale_shape_manual(values = c(1, 16)) +
  scale_linetype_manual(values = c("solid", "dashed"))

@@

max_pupil |>
  dplyr::mutate(trial_c = trial_c + 1) |>
  dplyr::group_by(speaker, rep_acc, trial_c) |>
  dplyr::summarize(peak = mean(peak_pupil, na.rm = T), 
                   se = sd(peak_pupil, na.rm = T)/sqrt(n())) |>
  ggplot() +
  aes(x = trial_c,
      y = peak,
      color = speaker,
      fill = speaker,
      linetype = speaker,
      shape = speaker) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Trial", y = "Peak Pupil Dilation (Arbitrary Units)") +
  facet_wrap("rep_acc") +
  theme_bw() +
  theme(legend.position = "bottom",
        aspect.ratio = 1) +
  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
  scale_fill_manual(values = c(my_pal[1], my_pal[5])) +
  scale_shape_manual(values = c(1, 16)) +
  scale_linetype_manual(values = c("solid", "dashed"))

@@

df_pred |>
  dplyr::mutate(condition = factor(condition, levels = c("Control Accurate", "Control Inaccurate",
                                                         "ALS Accurate", "ALS Inaccurate"))) |>
  dplyr::filter(time_norm >= 0 & time_norm <= 5000) |>
  ggplot() +
  aes(x = time_norm,
      y = trial_c,
      group = condition) +
  geom_tile(aes(fill = fit)) +
  geom_contour(aes(z = fit), colour = "white") +
  facet_wrap("condition") +
  theme_classic() +
  labs(fill = "Model Estimate",
       x = "Normed Time (ms)",
       y = "Trial Number") +
  scale_fill_viridis(option = "D") +
  theme(aspect.ratio = 1,
        legend.position = "bottom",
        axis.line = element_blank(),
        axis.text.x = element_text(angle = 45, hjust=1))

ggsave("Figures/pupil_heatmap.png", plot = last_plot())

>>>>

rep_acc_df |>
  dplyr::mutate(trial_c = trial_c + 1) |>
  dplyr::group_by(speaker, trial_c, rep_acc) |>
  dplyr::summarize(n = n()) |>
  dplyr::ungroup() |>
  tidyr::pivot_wider(names_from = "rep_acc",
                     values_from = "n") |>
  dplyr::mutate(accurate = ifelse(is.na(accurate), 0, accurate),
                inaccurate = ifelse(is.na(inaccurate), 0, inaccurate),
                prop_accurate = accurate/(accurate + inaccurate)) |>
  ggplot() +
  aes(x = trial_c,
      y = prop_accurate,
      group = speaker,
      color = speaker,
      fill = speaker, 
      linetype = speaker,
      shape = speaker) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(y = "Proportion of Accurate Responses",
       x = "Trial Number",
       color = "Speaker", fill = "Speaker", 
       linetype = "Speaker", shape = "Speaker") +
  theme_bw() +
  theme(legend.position = "bottom",
        aspect.ratio = 1) +
  scale_color_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
  scale_fill_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
  scale_shape_manual(values = c(1, 16), labels = c("Neurotypical", "ALS")) +
  scale_linetype_manual(values = c("solid", "dashed"), labels = c("Neurotypical", "ALS"))

@@

ple_data |>
  dplyr::mutate(trial_c = trial_c + 1) |>
  dplyr::group_by(speaker, rep_acc, trial_c) |>
  dplyr::summarize(per_effort = mean(effort_rating, na.rm = T), 
                   sd = sd(effort_rating, na.rm = T),
                   se = sd/sqrt(n())) |>
  ggplot() +
   aes(x = trial_c,
       y = per_effort,
       group = speaker,
       color = speaker,
       fill = speaker,
       shape = speaker,
       linetype = speaker) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Trial", y = "Perceived Listening Effort Rating",
       color = "Speaker", fill = "Speaker",
       shape = "Speaker", linetype = "Speaker") +
  facet_wrap("rep_acc") +
  theme_bw() +
  theme(legend.position = "bottom",
        aspect.ratio = 1) +
  scale_color_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
  scale_fill_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
  scale_shape_manual(values = c(1, 16), labels = c("Neurotypical", "ALS")) +
  scale_linetype_manual(values = c("solid", "dashed"), labels = c("Neurotypical", "ALS"))

@@

max_pupil |>
  dplyr::mutate(trial_c = trial_c + 1) |>
  dplyr::group_by(speaker, rep_acc, trial_c) |>
  dplyr::summarize(peak = mean(peak_pupil, na.rm = T), 
                   se = sd(peak_pupil, na.rm = T)/sqrt(n())) |>
  ggplot() +
  aes(x = trial_c,
      y = peak,
      color = speaker,
      fill = speaker,
      linetype = speaker,
      shape = speaker) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Trial", y = "Peak Pupil Dilation (Arbitrary Units)",
       color = "Speaker", fill = "Speaker", linetype = "Speaker", shape = "Speaker") +
  facet_wrap("rep_acc") +
  theme_bw() +
  theme(legend.position = "bottom",
        aspect.ratio = 1) +
  scale_color_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
  scale_fill_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
  scale_shape_manual(values = c(1, 16), labels = c("Neurotypical", "ALS")) +
  scale_linetype_manual(values = c("solid", "dashed"), labels = c("Neurotypical", "ALS"))

@@

df_pred |>
  dplyr::mutate(condition = factor(condition, levels = c("Control Accurate", "Control Inaccurate",
                                                         "ALS Accurate", "ALS Inaccurate"),
                                   labels = c("Neurotypical Accurate", "Neurotypical Inaccurate",
                                                         "ALS Accurate", "ALS Inaccurate"))) |>
  dplyr::filter(time_norm >= 0 & time_norm <= 5000) |>
  ggplot() +
  aes(x = time_norm,
      y = trial_c,
      group = condition) +
  geom_tile(aes(fill = fit)) +
  geom_contour(aes(z = fit), colour = "white") +
  facet_wrap("condition") +
  theme_classic() +
  labs(fill = "Pupil Dilation Estimate",
       x = "Normed Time (ms)",
       y = "Trial Number") +
  scale_fill_viridis(option = "D") +
  theme(aspect.ratio = 1,
        legend.position = "bottom",
        axis.line = element_blank(),
        axis.text.x = element_text(angle = 45, hjust=1))

ggsave("Figures/pupil_heatmap.png", plot = last_plot(), scale = 1.8)

### 2024-05-29 Manuscript Analysis/main_analysis.Rmd
updated figures

ple_data |>
  dplyr::group_by(speaker, rep_acc) |>
  dplyr::summarize(per_effort = mean(effort_rating, na.rm = T), 
                   sd = sd(effort_rating, na.rm = T),
                   se = sd/sqrt(n())) |>
  ggplot() +
   aes(x = rep_acc,
       y = per_effort,
       group = speaker,
       color = speaker,
       fill = speaker) +
  geom_bar(stat = "identity", alpha = 0.6, position = position_dodge()) +
  geom_errorbar(aes(ymin = per_effort - se, ymax = per_effort + se), width = 0.4, position = position_dodge(.9)) +
  labs(x = "Accuracy", y = "Perceived Listening Effort Rating") +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
  scale_fill_manual(values = c(my_pal[1], my_pal[5]))

@@

m2_predicted_values |>
  ggplot() +
  aes(x = time_norm,
      y = fit,
      color = speaker,
      fill = speaker,
      linetype = speaker) +
  geom_ribbon(aes(ymin = ul, ymax = ll), alpha = .5, color = NA) +
  geom_line() +
  geom_vline(xintercept = 0, size = 0.35) +
  geom_vline(xintercept = 1910, linetype = 2) +
  annotate("rect", xmin = 500, xmax = 2910, ymin = 0, ymax = 200, alpha = 0.15) +
  coord_cartesian(ylim = c(-10, 250), xlim = c(0, 4910)) +
  facet_wrap("rep_acc") +
  theme_bw() +
  theme(aspect.ratio = 1,
        legend.position = "bottom") +
  labs(x = "Normalized Time (ms)",
       y = "Pupil Dilation (Arbitrary Units)") +
  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
  scale_fill_manual(values = c(my_pal[1], my_pal[5])) +
  scale_linetype_manual(values = c("solid", "dashed"))

@@

ggsave("Figures/difference_plots_speakers.png", plot = diff_plot_1)

>>>>

ple_data |>
  dplyr::group_by(speaker, rep_acc) |>
  dplyr::summarize(per_effort = mean(effort_rating, na.rm = T), 
                   sd = sd(effort_rating, na.rm = T),
                   se = sd/sqrt(n())) |>
  ggplot() +
   aes(x = rep_acc,
       y = per_effort,
       group = speaker,
       color = speaker,
       fill = speaker) +
  geom_bar(stat = "identity", alpha = 0.6, position = position_dodge()) +
  geom_errorbar(aes(ymin = per_effort - se, ymax = per_effort + se), width = 0.4, position = position_dodge(.9)) +
  labs(x = "Accuracy", y = "Perceived Listening Effort Rating", color = "Speaker", fill = "Speaker") +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_color_manual(values = c(my_pal[1], my_pal[5]), labels =  c("Neurotypical", "ALS")) +
  scale_fill_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS"))

@@

m2_predicted_values |>
  ggplot() +
  aes(x = time_norm,
      y = fit,
      color = speaker,
      fill = speaker,
      linetype = speaker) +
  geom_ribbon(aes(ymin = ul, ymax = ll), alpha = .5, color = NA) +
  geom_line() +
  geom_vline(xintercept = 0, size = 0.35) +
  geom_vline(xintercept = 1910, linetype = 2) +
  annotate("rect", xmin = 500, xmax = 2910, ymin = 0, ymax = 200, alpha = 0.15) +
  coord_cartesian(ylim = c(-10, 250), xlim = c(0, 4910)) +
  facet_wrap("rep_acc") +
  theme_bw() +
  theme(aspect.ratio = 1,
        legend.position = "bottom") +
  labs(x = "Normalized Time (ms)",
       y = "Pupil Dilation (Arbitrary Units)",
       color = "Speaker",
       fill = "Speaker",
       linetype = "Speaker") +
  scale_color_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
  scale_fill_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
  scale_linetype_manual(values = c("solid", "dashed"), labels = c("Neurotypical", "ALS"))

@@

ggsave("Figures/difference_plots_speakers.png", plot = diff_plot_1, scale = 1.8)
